---
title: "w203_lab2: regression models draft"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("stargazer")
library(stargazer)
library(tidyverse)
library(sandwich)
```

```{r data schema}
schema <- cols(
  state = "c",
  cases_total = "i",
  cases_last_7_days = "i",
  case_rate = "n",
  case_rate_last_7_days = "n",
  deaths_total = "i",
  deaths_last_7_days = "i",
  death_rate = "n",
  death_rate_last_7_days = "n",
  tests_total = "i",
  tests_positive = col_factor(
    levels = c("0-5%", "6-10%", "11-20%"),
    ordered = TRUE
    ),
  test_rate = "i",
  white_cases = "i",
  white_pop = "i",
  black_cases = "i",
  black_pop = "i",
  hispanic_cases = "i",
  hispanic_pop = "i",
  other_cases = "i",
  other_pop = "i",
  white_deaths = "i",
  black_deaths = "i",
  hispanic_deaths = "i",
  other_deaths = "i",
  emerg_date = col_date(format = "%d/%m/%Y"),
  beg_bus_close_date = col_date(format = "%d/%m/%Y"),
  end_bus_close_date = col_date(format = "%d/%m/%Y"),
  bus_close_days = "i",
  beg_shelter_date = col_date(format = "%d/%m/%Y"),
  end_shelter_date = col_date(format = "%d/%m/%Y"),
  shelter_days = "i",
  mask_date = col_date(format = "%d/%m/%Y"),
  mask_use = "l",
  mask_legal = "l",
  beg_maskbus_date = col_date(format = "%d/%m/%Y"),
  end_maskbus_date = col_date(format = "%d/%m/%Y"),
  maskbus_use = "l",
  gov_party = "l",
  pop_dens = "n",
  pop_total = "i",
  pre_cond_total = "i",
  serious_illness_pct = "n",
  all_cause_deaths_total = "i",
  homeless_total = "i",
  medicaid_pct = "i",
  life_expectancy = "n",
  unemployment_rate = "n",
  poverty_rate = "n",
  weekly_UI_max_amount = "i",
  household_income = "i",
  age_0_18 = "i",
  age_19_25 = "i",
  age_26_34 = "i",
  age_35_54 = "i",
  age_55_64 = "i",
  age_65 = "i",
  mob_RR = "i",
  mob_GP = "i",
  mob_PK = "i",
  mob_TS = "i",
  mob_WP = "i",
  mob_RS = "i"
)
```

```{r csv import}
df <- read_delim(
  file = "clean_covid_19_LB_version.csv",
  delim = ";",
  col_names = TRUE,
  col_types = schema,
  na = ""
  )
```

Question: Should we include test_rate (or any transformation of it) as an initial variable on our model?

Answer: Yes, we should include sqrt_test_rate on our initial model version

```{r test_rate linear relationship}
plot1 <- df %>%
  ggplot(aes(y = case_rate, x = test_rate)) +
  geom_point() +
  labs(
    title = "Relationship between Test Rate and Case Rate by state",
    x = "Test Rate per 100K",
    y = "Case Rate per 100K"
  )
```

```{r test_rate squared relationship}
plot2 <- df %>%
  ggplot(aes(y = case_rate, x = test_rate^2)) +
  geom_point() +
  labs(
    title = "Relationship between Squared Test Rate and Case Rate by state",
    x = "Squared Test Rate per 100K",
    y = "Case Rate per 100K"
  )
```

```{r creating sqrt_test_rate}
df$sqrt_test_rate = df$test_rate^2
```

```{r alternatives for model version #1}

mod1_1 <- lm(case_rate ~ 
             mask_use, 
           data = df
           )

mod1_2 <- lm(case_rate ~
             mask_use +
             test_rate,
           data = df
           )

mod1_3 <- lm(case_rate ~
             mask_use +
             sqrt_test_rate,
           data = df
           )

std_errors = list(
  sqrt(diag(vcovHC(mod1_1))), 
  sqrt(diag(vcovHC(mod1_2))), 
  sqrt(diag(vcovHC(mod1_3)))
  )

stargazer(mod1_1, mod1_2, mod1_3, se = std_errors, type = "text")
```

Question: Should we include pop_dens as an another initial variable on our model on top of test_rate?

Answer: No, we should not add pop_dens to our regression model

```{r pop_dens linear relationship}
plot3 <- df %>%
  ggplot(aes(y = case_rate, x = pop_dens)) +
  geom_point() +
  labs(
    title = "Relationship between Population Density and Case Rate by state",
    x = "Population Density",
    y = "Case Rate per 100K"
  )
```

```{r pop_dens log relationship}
plot4 <- df %>%
  ggplot(aes(y = case_rate, x = log(pop_dens))) +
  geom_point() +
  labs(
    title = "Relationship between Log Population Density and Case Rate by state",
    x = "Log Population Density",
    y = "Case Rate per 100K"
  )
```

```{r alternatives for model version #2}

mod2_1 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate,
              data = df
                )

mod2_2 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate +
                pop_dens,
              data = df
                )

mod2_3 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate +
                log(pop_dens),
              data = df
                )

std_errors = list(
  sqrt(diag(vcovHC(mod2_1))), 
  sqrt(diag(vcovHC(mod2_2))), 
  sqrt(diag(vcovHC(mod2_3)))
  )

stargazer(mod2_1, mod2_2, mod2_3, se = std_errors, type = "text")
```

Question: Should we include any variable to control for age demographics? If yes, which variable does the better job in improving our model explanability?

Answer: Yes, we should include age_below_25

```{r variance-covariance matrix of case_rate vs. age variables}
var(df[ ,c(4, 51:56)])
```

```{r creating age_below_25 and age_above_55}
df$age_below_25 = df$age_0_18 + df$age_19_25
df$age_above_55 = df$age_55_64 + df$age_65
```

```{r variance-covariance matrix or case_rate vs. new age variables}
var(df[ ,c(4, 64:65)])
```

```{r age_below_25 scatterplot}
plot5 <- df %>%
  ggplot(aes(y = case_rate, x = age_below_25)) +
  geom_point() +
  labs(
    title = "Relationship between Pct Below 25 and Case Rate by state",
    x = "% of Population Below 25",
    y = "Case Rate per 100K"
  )
```

```{r age_above_55 scatterplot}
plot6 <- df %>%
  ggplot(aes(y = case_rate, x = age_above_55)) +
  geom_point() +
  labs(
    title = "Relationship between Pct Above 55 and Case Rate by state",
    x = "% of Population Above 55",
    y = "Case Rate per 100K"
  )
```

```{r alternatives for model version #3}
mod3_1 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate,
              data = df
                )

mod3_2 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate +
                age_below_25,
              data = df
                )

mod3_3 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate +
                age_above_55,
              data = df
                )

std_errors = list(
  sqrt(diag(vcovHC(mod3_1))), 
  sqrt(diag(vcovHC(mod3_2))), 
  sqrt(diag(vcovHC(mod3_3)))
  )

stargazer(mod3_1, mod3_2, mod3_3, se = std_errors, type = "text")
```

Question: Should we include any variable to control for socio-economic differences among states? If yes, which variable does the better job in improving our model explanability?

Answer: No, we should not include any variable to count for socio-economic differences

```{r variance-covariance matrix of case_rate vs. socio-economic variables}
var(df[ ,c(4, 44, 47, 48, 50)], na.rm = TRUE)
```

```{r homeless_total scatterplot}
plot7 <- df %>%
  ggplot(aes(y = case_rate, x = homeless_total)) +
  geom_point() +
  labs(
    title = "Relationship between Homeless Population and Case Rate by state",
    x = "Homeless Population",
    y = "Case Rate per 100K"
  )
```

```{r log homeless_total scatterplot}
plot8 <- df %>%
  ggplot(aes(y = case_rate, x = log(homeless_total))) +
  geom_point() +
  labs(
    title = "Relationship between Log Homeless Population and Case Rate by state",
    x = "Log Homeless Population",
    y = "Case Rate per 100K"
  )
```

```{r household income scatterplot, warning = FALSE}
plot9 <- df %>%
  ggplot(aes(y = case_rate, x = household_income)) +
  geom_point() +
  labs(
    title = "Relationship between Median Household Income and Case Rate by state",
    x = "Household Income",
    y = "Case Rate per 100K"
  )
```

```{r alternatives for model version #4}
mod4_1 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate +
                age_below_25,
              data = df
                )

mod4_2 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate +
                age_below_25 +
                log(homeless_total),
              data = df
                )

mod4_3 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate +
                age_below_25 +
                household_income,
              data = df
                )

std_errors = list(
  sqrt(diag(vcovHC(mod4_1))), 
  sqrt(diag(vcovHC(mod4_2))), 
  sqrt(diag(vcovHC(mod4_3)))
  )

stargazer(mod4_1, mod4_2, mod4_3, se = std_errors, type = "text")
```

Question: Should we include any variable to control for race mix differences among states? If yes, which variable does the better job in improving our model explanability?

Answer: Yes, we should include the log(black_pop)

```{r variance-covariance matrix of case_rate vs. race variables}
var(df[ ,c(4, 14, 16, 18)])
```

```{r black_pop scatterplot}
plot10 <- df %>%
  ggplot(aes(y = case_rate, x = black_pop)) +
  geom_point() +
  labs(
    title = "Relationship between % of Blacks and Case Rate by state",
    x = "% of Blacks in Total Population",
    y = "Case Rate per 100K"
  )
```

```{r log black_pop scatterplot}
plot11 <- df %>%
  ggplot(aes(y = case_rate, x = log(black_pop))) +
  geom_point() +
  labs(
    title = "Relationship between Log % of Blacks and Case Rate by state",
    x = "Log % of Blacks in Total Population",
    y = "Case Rate per 100K"
  )
```

```{r dealing with zeros on black_pop}
df$black_pop[df$black_pop == 0] = 0.01
```

```{r lternatives for model version #5}
mod5_1 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate +
                age_below_25,
              data = df
                )

mod5_2 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate +
                age_below_25 +
                black_pop,
              data = df
                )

mod5_3 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate +
                age_below_25 +
                log(black_pop),
              data = df
                )

std_errors = list(
  sqrt(diag(vcovHC(mod5_1))), 
  sqrt(diag(vcovHC(mod5_2))), 
  sqrt(diag(vcovHC(mod5_3)))
  )

stargazer(mod5_1, mod5_2, mod5_3, se = std_errors, type = "text")
```

Question: Should we include any indicator from Google mobility? If yes, which variable does the better job in improving our model explanability?

Answer: Yes, we should include the mob_TS variable 

```{r variance-covariance matrix of case_rate vs. mobility variables, warning=FALSE}
var(df[ ,c(4, 57:62)])
```

```{r mob_TS scatterplot}
plot12 <- df %>%
  ggplot(aes(y = case_rate, x = mob_TS)) +
  geom_point() +
  labs(
    title = "Relationship between Mobility in Terminal Stations and Case Rate by state",
    x = "Mobility in Terminal Stations",
    y = "Case Rate per 100K"
  )
```

```{r alternatives for model version #6, warning=FALSE}
mod6_1 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate +
                age_below_25 +
                log(black_pop),
              data = df
                )

mod6_2 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate +
                age_below_25 +
                log(black_pop) +
                mob_TS,
              data = df
                )

std_errors = list(
  sqrt(diag(vcovHC(mod6_1))), 
  sqrt(diag(vcovHC(mod6_2)))
  )

stargazer(mod6_1, mod6_2, se = std_errors, type = "text")
```

Question: Should we include any other variable related to policies adopted by states? If yes, which variable does the better job on improving our model explanability?

Answer: No, we should not include any variable related to other policies

```{r variance-covariance matrix of case_rate vs. policies variables, warning=FALSE}
var(df[,c(4, 28, 31, 34, 37, 38)], na.rm=TRUE)
```

```{r shelter_days scatteplot, warning=FALSE}
plot13 <- df %>%
  ggplot(aes(y = case_rate, x = shelter_days)) +
  geom_point() +
  labs(
    title = "Relationship between Shelter in Place and Case Rate by state",
    x = "Number of Days of Shelter in Place",
    y = "Case Rate per 100K"
  )
```

```{r alternatives for model version #7, warning=FALSE}
mod7_1 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate +
                age_below_25 +
                log(black_pop) +
                mob_TS,
              data = df
                )

mod7_2 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate +
                age_below_25 +
                log(black_pop) +
                mob_TS +
                shelter_days,
              data = df
                )

std_errors = list(
  sqrt(diag(vcovHC(mod7_1))), 
  sqrt(diag(vcovHC(mod7_2)))
  )

stargazer(mod7_1, mod7_2, se = std_errors, type = "text")
```

Question: What should be our final three model versions?

Answer:
model_1 ~ mask_use + sqrt_test_rate
model_2 ~ mask_use + sqrt_test_rate + below_25 + log(black_pop)
model_3 ~ mask_use + sqrt_test_rate + below_25 + log(black_pop) + shelter_days + bus_close_days

model_1 is point of departure
model_2 is our best model
model_3 is aimed to stress the significance of our coefficient when we add another policies that compete for variability with mask_use


```{r, warning=FALSE}
mod8_1 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate,
              data = df
                )

mod8_2 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate +
                age_below_25 +
                log(black_pop) +
                mob_TS,
              data = df
                )

mod8_3 <- lm (case_rate ~
                mask_use +
                sqrt_test_rate +
                age_below_25 +
                log(black_pop) +
                mob_TS +
                shelter_days +
                bus_close_days,
              data = df
                )

std_errors = list(
  sqrt(diag(vcovHC(mod8_1))), 
  sqrt(diag(vcovHC(mod8_2))),
  sqrt(diag(vcovHC(mod8_3)))
  )

stargazer(mod8_1, mod8_2, mod8_3, se = std_errors, type = "text")
```

