---
title: "R Notebook"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("stargazer")
library(stargazer)
library(tidyverse)
library(sandwich)
```

```{r data schema, warning=FALSE}
schema <- cols(
  state = "c",
  cases_total = "i",
  cases_last_7_days = "i",
  case_rate = "n",
  case_rate_last_7_days = "n",
  deaths_total = "i",
  deaths_last_7_days = "i",
  death_rate = "n",
  death_rate_last_7_days = "n",
  tests_total = "i",
  tests_positive = col_factor(
    levels = c("0-5%", "6-10%", "11-20%"),
    ordered = TRUE
    ),
  test_rate = "i",
  white_cases = "i",
  white_pop = "i",
  black_cases = "i",
  black_pop = "i",
  hispanic_cases = "i",
  hispanic_pop = "i",
  other_cases = "i",
  other_pop = "i",
  white_deaths = "i",
  black_deaths = "i",
  hispanic_deaths = "i",
  other_deaths = "i",
  emerg_date = col_date(format = "%d/%m/%Y"),
  beg_bus_close_date = col_date(format = "%d/%m/%Y"),
  end_bus_close_date = col_date(format = "%d/%m/%Y"),
  bus_close_days = "i",
  beg_shelter_date = col_date(format = "%d/%m/%Y"),
  end_shelter_date = col_date(format = "%d/%m/%Y"),
  shelter_days = "i",
  mask_date = col_date(format = "%d/%m/%Y"),
  mask_use = "i",
  mask_legal = "l",
  beg_maskbus_date = col_date(format = "%d/%m/%Y"),
  end_maskbus_date = col_date(format = "%d/%m/%Y"),
  maskbus_use = "l",
  gov_party = col_factor(
    levels = c("R", "D"),
    ordered = FALSE
  ),
  pop_dens = "n",
  pop_total = "i",
  pre_cond_total = "i",
  serious_illness_pct = "n",
  all_cause_deaths_total = "i",
  homeless_total = "i",
  medicaid_pct = "i",
  life_expectancy = "n",
  unemployment_rate = "n",
  poverty_rate = "n",
  weekly_UI_max_amount = "i",
  household_income = "i",
  age_0_18 = "i",
  age_19_25 = "i",
  age_26_34 = "i",
  age_35_54 = "i",
  age_55_64 = "i",
  age_65 = "i",
  mob_RR = "i",
  mob_GP = "i",
  mob_PK = "i",
  mob_TS = "i",
  mob_WP = "i",
  mob_RS = "i"
)
```

```{r csv import, warning=FALSE}
df <- read_delim(
  file = "clean_covid_19_LB_version.csv",
  delim = ";",
  col_names = TRUE,
  col_types = schema,
  na = ""
  )
```

```{r}
df = mutate(df, percentage_homeless = (homeless_total/pop_total)*100)
```

### Omitted Variables Discussion

As we are attempting to study an exceedingly-complex real life phenomenon, naturally there will be an element of bias in our models. Even though we selected what we believed to be the most appropriate variables influencing the COVID-19 case rate per 100,000, which included a range of logistical, social, ethnic and population-based metrics (as provided in the dataset), it is impossible to completely predict the influences on our dependent variable - this is reflected in the model metrics, for example R2, which would be 1 if we could perfectly know and include all of the real-life variables contributing to the case rate per 100,000.  

As such, we have selected several omitted variables that we believe to contribute to the COVID-19 test rate. Some of these are included in the dataset, while others are phenomena that we do not have any information regarding. Though we can predict the potential relationship between the omitted variables and the dependent and independent variables for all of our suggested omitted variables, we can only test the veracity of our predictions for the former. We discuss the omitted variables in relation to our "most complete" models, i.e. ones with what we believe to be the optimal number of explanatory variables. 

The seven independent variables included in this model are **Mandatory Mask Use**, **Test Rate per 100,000 habitants**, **Percentage of Population Below 25 Years Old**, **Log of Percentage of Black Ethnicity in Total Population**, and **Human Mobility Change in Transit Stations**, which have all been discussed in relation to homelessness as an omitted variable, previously. Additionally we include  **Number of Days of Shelter in Place**, and **Number of Days of Non-Essential Businesses Closure** in our model. 

## Percentage of Homeless Population per State

This variable is included in the dataset. 

The percentage of homeless people per state population can broadly be categorized as a socioeconomic metric. Due to their circumstances, homeless people are less likely to be able to afford PPE, and have limited access to accommodation that suitably limits their exposure to others (if they have any accommodation at all). This theory is largely supported by [existing research](https://www.thelancet.com/journals/lanres/article/PIIS2213-2600(20)30396-9/fulltext), and consequently there have even been initiatives to provide temporary housing in an effort to control the spread of the infection among this group and the larger population. Nevertheless, these initiatives are inconsistent across regions, and in regions where they have been offered, [they have been unsuccessful](https://www.latimes.com/homeless-housing/story/2020-05-19/newsom-coronavirus-homeless-population-hotel-project-roomkey).

As such we believe there is a positive correlation between homelessness and the dependent variable, i.e. COVID-19 case rate per 100,000. That is, in States with higher percentages of homeless people, the case rate per 100,000 will be higher. 

**Mandatory Mask Use**

According to this model, the enforcement of mask usage reduces the case rate by ~941 cases/100,000, holding the other variable constant.

States with higher numbers of homeless populations in theory should be more aware of the risks of COVID-19 spreading among this vulnerable population and will therefore be more likely to be proactive in terms of public health measures, including mandating mask use. for the purposes of this exercise, we will assume that there is a positive relationship between mask use enforcement and level of homelessness.   

Given that we predict a positive relationship between homelessness and the dependent variable (case rate per 100,000), and a positive relationship between homelessness and mask use enforcement, the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the homelessness variable in the model, it will cause the mask use coefficient to tend towards 0, or become less significant. As the coefficient is already negative, we would expect to see it becoming less negative.  

**Test Rate per 100,000 habitants**

According to the results of the model, increasing the test rate per 100,000 by 1 unit will increase the number of positive tests per 100,000 by 0.023. 

Most States have moved to provide no-cost COVID-19 testing at health centers and pharmacies. Moreover, as with mask use, we assume that States with large homeless populations are also more proactive in terms of public health measures, including COVID-19 testing. We therefore predict that there is a positive relationship between the percentage of homeless people in a state and the test rate per 100,000. 

Given that we predict a positive relationship between homelessness and the dependent variable (case rate per 100,000) and a positive relationship between homelessness and test rate per 100,000 people, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the homelessness variable in the model, it will cause the test rate coefficient to tend towards 0, or become less significant. 

**Percentage of Population Below 25 Years Old**

According to the results of the model, a one unit increase in the percentage of the population under the age of 25 increases the positive case rate by ~190 per 100,000 of the population, holding all other variables constant. 

Looking at maps of distributions of each metric in the United States, [the distribution of young people, i.e. < 25 YO](https://www.kff.org/other/state-indicator/distribution-by-age/?activeTab=map&currentTimeframe=0&selectedDistributions=children-0-18&sortModel=%7B%22colId%22:%22Location%22,%22sort%22:%22asc%22%7D) seems to be concentrated primarily within the non-Coastal areas, whereas the coastal areas appear to have a [higher concentration of homeless people](https://www.usich.gov/tools-for-action/map/#fn[]=1400&fn[]=2800&fn[]=6200&fn[]=10000&fn[]=13200). Using these approximate trends, we propose that there is a negative relationship between homelessness and age. 

Given that we predict a positive relationship between homelessness and the dependent variable (case rate per 100,000), and a negative relationship between homelessness and percentage of young people in the State, we predict that the overall predicted variable bias (disregarding all other independent variables) will be negative. That is, if we include the homelessness variable in the model, it will cause the percentage under 25 variable to tend away from zero, or become more significant. 

**Log of Percentage of Black Ethnicity in Total Population**

According to the results of the model, a one unit increase in the log of the percentage of black ethnicity in the population causes the rate of positive tests to increase by ~220 cases per 100,000 population. 

Black people, despite comprising 13% of the U.S. population, [account for 40% of the homeless](https://endhomelessness.org/homelessness-in-america/what-causes-homelessness/inequality/). The reasons for this are well known and tied to complex socioeconomic factors. Subsequently, we predict a positive relationship between the percentage of black people within a state and the state percentage of homeless people. 

Given that we predict a positive relationship between homelessness and the dependent variable (case rate per 100,000), and a positive relationship between homelessness and the percentage of black people within a population, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the homelessness variable in the model, it will cause the log of the percentage of black people coefficient to tend towards 0, or become less significant. 

**Human Mobility Change in Transit Stations**

According to the results of the model a one unit increase in the mobility change in transit stations variable causes the rate of positive tests to increase by ~16 cases per 100,000 population. 

Transit stations are [well known congregation areas](https://www.latimes.com/local/lanow/la-me-metro-homeless-20180406-htmlstory.html) for homeless people. They may seek shelter or panhandle. As the pandemic has largely disrupted peoples work routines, with many now working from home, transit stations may have become less busy and may offer more space for homeless people to congregate or seek shelter. As a result, we predict a negative relationship between the human mobility change and homelessness variables.

Given that we predict a positive relationship between homelessness and the dependent variable (case rate per 100,000), and a negative relationship between homelessness and mobility change in transit stations, we predict that the overall predicted variable bias (disregarding all other independent variables) will be negative. That is, if we include the homelessness variable in the model, it will cause the human mobility change variable to tend away from zero, or become more significant.

**Number of Days of Shelter in Place**

According to the results of the model a one unit increase in the number of shelter in place days reduces the case rate per 100,000 by ~0.8.

As discussed in percentage of population below 25 discussion, the homeless population seems to be largely concentrated in the coastal areas. The coastal cities, being more metropolitan and having a higher proportion of educated population were likely to enforce evidence-based doctrines recommending institution of shelter in place rules earlier, relative to the non-coastal US states. As such, we predict there to be a positive correlation between percent of homeless population and the number of days of shelter in place. 

Given that we predict a positive relationship between homelessness and the dependent variable (case rate per 100,000), and a positive relationship between homelessness and the number of days of shelter in place, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the homelessness variable in the model, it will cause the shelter days coefficient to tend towards 0, or become less significant. As the coefficient is already negative, we would expect to see it becoming less negative.

**Number of Days of Non-Essential Businesses Closure**

According to the results of the model a one unit increase in the number of days that non-essential businesses remained closed causes the case rate to increase by ~6.9 cases per 100,000. 

For the same reasons that we discussed for the previous variable regarding shelter-in-place we predict a positive relationship between homeless population and number of business closure days. States more likely to have imposed strict rules regarding business and activity are more likely to be the ones with high levels of homeless population. 

Given that we predict a positive relationship between homelessness and the dependent variable (case rate per 100,000), and a positive relationship between homelessness and the number of days of forced essential business closure, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the homelessness variable in the model, it will cause the business closure days coefficient to tend towards 0, or become less significant. 

## Percentage of Hispanic Population per State

This variable is included in the dataset. 

People of Hispanic descent comprise 17% of the U.S. population, however their distribution varies by state, with the population ranging from [30-50% in some Southern and South Western States, to less than 5% in Mississipi and West Virginia](https://en.wikipedia.org/wiki/Demographics_of_Hispanic_and_Latino_Americans). Moreover, approximately [19% of Hispanic origin people are estimated to live in poverty](https://www.pewresearch.org/hispanic/fact-sheet/u-s-hispanics-facts-on-spanish-origin-latinos/#:~:text=%2450%2C000).-,Poverty%20status,%25%20of%20foreign%2Dborn%20Spaniards.). Together with black people, Hispanics comprise [55% of COVID-19 cases](https://salud-america.org/55-percent-of-u-s-coronavirus-cases-are-latino-black/). For these socioeconomic and demographic reasons, we believe that there is a positive relationship between the states with high percentage population of Hispanics and the number of positive COVID-19 cases per 100,000 population. 

**Mandatory Mask Use**

According to this model, the enforcement of mask usage reduces the case rate by ~941 cases/100,000, holding the other variable constant.

Though [California](https://calmatters.org/explainers/coronavirus-california-explained-newsom/) and [New Mexico](https://www.scientificamerican.com/article/how-new-mexico-controlled-the-spread-of-covid-19/) States with large Hispanic diasporas implemented early evidence-based measures (including mandatory mask usage) to stem the spread of COVID-19. New York, another State with a large Hispanic population became the National epicenter for coronavirus relatively early on in the pandemic and as a result issued mandatory mask orders early-on. As such, we predict a positive correlation between the States that imposed mandatory mask wearing in public areas and the percentage of Hispanic population. 

Given that we predict a positive relationship between Hispanic population and the dependent variable (case rate per 100,000), and a positive relationship between Hispanic population and mask use enforcement per state, the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the percentage of Hispanic population variable in the model, it will cause the mask use coefficient to tend towards 0, or become less significant. As the coefficient is already negative, we would expect to see it becoming less negative. 

**Test Rate per 100,000 habitants**

According to this model, increasing the test rate per 100,000 by 1 unit will increase the number of positive tests per 100,000 by 0.023.

Hispanics are often employed as [essential workers](https://abcnews.go.com/US/latinos-essential-jobs-disproportionately-affected-covid-19/story?id=71928362), in positions that may make them increasingly susceptible to COVID-19. We presume that States with high Hispanic populations have conducted appropriate public health initiatives to educate the Hispanic population in regards to COVID-19 testing (e.g. Spanish-language materials etc.), thus we predict a positive relationship between the population of Hispanics and test rate per 100,000 inhabitants. 

Given that we predict a positive relationship between homelessness and the dependent variable (case rate per 100,000) and a positive relationship between test rate and population of Hispanic people, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the Hispanic population variable in the model, it will cause the test rate coefficient to tend towards 0, or become less significant.

**Percentage of Population Below 25 Years Old**

According to the results of the model, a one unit increase in the percentage of the population under the age of 25 increases the positive case rate by ~190 per 100,000 of the population, holding all other variables constant.

[Approximately one third of the Spanish population in the U.S. is under the age of 18, while one quarter are millenials](https://www.pewresearch.org/hispanic/2016/04/20/the-nations-latino-population-is-defined-by-its-youth/). Due to the high proportion of young people, we predict a positive relationship between Hispanic population and percentage of population under age 25, per State. 

Given that we predict a positive relationship between Hispanic population and the dependent variable (case rate per 100,000), and a positive relationship between Hispanic population and percentage of young people in the State, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the Hispanic population variable in the model, it will cause the percentage  coefficient to tend towards 0, or become less significant.

**Log of Percentage of Black Ethnicity in Total Population**

According to the results of the model, a one unit increase in the log of the percentage of black ethnicity in the population causes the rate of positive tests to increase by ~220 cases per 100,000 population.

Looking at maps of distributions of [Hispanics](https://en.wikipedia.org/wiki/List_of_U.S._states_by_Hispanic_and_Latino_population#/media/File:Percent_of_Hispanic_and_Latino_population_by_state_in_2012.svg) compared with [Black ethnicity](https://en.wikipedia.org/wiki/List_of_U.S._states_and_territories_by_African-American_population#/media/File:African_American_by_state_in_the_USA_in_2010.svg), there seems to be a negative correlation, i.e. States with higher concentrations of Hispanics have lower concentrations of Black people and vice versa. 

Given that we predict a postive relationship between Hispanic population and the dependent variable (case rate per 100,000), and a negative relationship between Hispanic people and the percentage of black people within a population, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be negative. That is, if we include the Hispanic population variable in the model, it will cause the log of Black ethnicity variable to tend away from zero, or become more significant.

**Human Mobility Change in Transit Stations**

According to the results of the model a one unit increase in the mobility change in transit stations variable causes the rate of positive tests to increase by ~16 cases per 100,000 population. 

- How to relate hispanics with this?

**Number of Days of Shelter in Place**

According to the results of the model a one unit increase in the number of shelter in place days reduces the case rate per 100,000 by ~0.8.

The three States with the longest running shelter-in-place orders (i.e. the ones that implemented it earliest) are California, New York and New Mexico. These three States also have some of the highest proportions of Hispanics in the country. We can take this as a general trend to predict that there is a positive relationship between the percentage of Hispanics in the population and number of shelter-in-place days. 

Given that we predict a positive relationship between Hispanic population percentage and the dependent variable (case rate per 100,000), and a positive relationship between Hispanic population percentage and the number of days of shelter in place, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the Hispanic population percentage variable in the model, it will cause the shelter days coefficient to tend towards 0, or become less significant. As the coefficient is already negative, we would expect to see it becoming less negative.

**Number of Days of Non-Essential Businesses Closure**

According to the results of the model a one unit increase in the number of days that non-essential businesses remained closed causes the case rate to increase by ~6.9 cases per 100,000.

It is likely that States concurrently issued shelter-in-place orders together with enforced closure of non-essential businesses. These States happened to also be ones with large Hispanic populations. Likewise, we also predict a positive relationship between number of days of non-essential business closure and percentage Hispanic population. 

Given that we predict a positive relationship between Hispanic population percentage and the dependent variable (case rate per 100,000), and a positive relationship between Hispanic population percentage and the number of days of forced essential business closure, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the homelessness variable in the model, it will cause the business closure days coefficient to tend towards 0, or become less significant. 

## Access to Healthcare

The United States is the only developed nation that does not offer some kind of free public healthcare option. Consequently 8.5% of the population or 27.5 million are without health insurance. This may be due to underemployment, poverty, disabilites, or a myriad of other reasons. Uninsured essential jobs may result in exposure to COVID, as may living in sub-standard accomodation. We therefore predict that reduced access to healthcare (higher percentage of uninsured population) results in a higher rate of positive COVID-19 tests per 100,000 of population (a positive relationship). As COVID-19 testing is largely free at the State-level we assume access to testing is unrelated to having health insurance.  

**Mandatory Mask Use**

According to this model, the enforcement of mask usage reduces the case rate by ~941 cases/100,000, holding the other variable constant. 

There are [16 States, all led by Republicans](https://undark.org/2020/10/06/16-states-dont-mandate-mask-use/) that have not implemented a mandatory mask order. Similarly, Republican-led States have also been hardest hit by the effects of globalization and a shifting National economy. These States therefore tend to have [higher rates of poverty](https://www.baltimoresun.com/opinion/editorial/bs-ed-0902-gop-states-underperform-20200831-kuquanxy4jd6zn55hqi76ery6i-story.html), and therefore will have higher percentage of uninsured individuals. We therefore predict that States with a higher percentage of uninsured in their populations will be less likely to enforce mandatory mask orders. 
Given that we predict a positive relationship between the percentage of uninsured and the dependent variable (case rate per 100,000), and a negative relationship between access to healthcare and mask use enforcement, the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be negative. That is, if we include a variable reflecting percentage uninsured, it will cause the mask use coefficient to tend away from 0, or become more significant. As the coefficient is already negative, we would expect to see it becoming more negative.  

**Test Rate per 100,000 habitants**

According to this model, increasing the test rate per 100,000 by 1 unit will increase the number of positive tests per 100,000 by 0.023.

For the same reasons discussed previously, we predict that States with higher proportions of uninsured may also be less likely to implement proactive public health policies, including expanding access to testing. We therefore predict a negative relationship between the test rate per 100,000 population and the percentage of uninsured individuals in the population. 

Given that we predict a positive relationship between the percentage of uninsured and the dependent variable (case rate per 100,000) and a negative relationship between percentage uninsured and test rate per 100,000 population, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be negative. That is, if we include the Hispanic population variable in the model, it will cause the test rate coefficient to tend away from 0, or become more significant.

**Percentage of Population Below 25 Years Old**

According to the results of the model, a one unit increase in the percentage of the population under the age of 25 increases the positive case rate by ~190 per 100,000 of the population, holding all other variables constant.

In the United States, there appears to be an inverse relationship between [household income and birth rate](https://www.statista.com/statistics/241530/birth-rate-by-family-income-in-the-us/). In 2017 those earning under $10,000 annually had the highest birth rates relative to other income groups. As low income is related to lack of healthcare access, we therefore predict that high proportions of uninsured individuals is positively related to percentage of State population under 25. 

Given that we predict a positive relationship between the percentage of uninsured and the dependent variable (case rate per 100,000), and a positive relationship between percentage uninsured and percentage of young people in the State, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the percentage uninsured variable in the model, it will cause the percentage  coefficient to tend towards 0, or become less significant.

**Log of Percentage of Black Ethnicity in Total Population**

According to the results of the model, a one unit increase in the log of the percentage of black ethnicity in the population causes the rate of positive tests to increase by ~220 cases per 100,000 population. 

Despite positive efforts at the Federal and State levels, Black people are [more likely](https://www.kff.org/racial-equity-and-health-policy/issue-brief/changes-in-health-coverage-by-race-and-ethnicity-since-the-aca-2010-2018/) to be uninsured relative to Caucasians and Asians. As a result, we predict a positive relationship between the percentage of Black people per state and the percentage of uninsured. 

Given that we predict a postive relationship between percentage of uninsured and the dependent variable (case rate per 100,000), and a positive relationship between the percentage of uninsured and the percentage of black people within a population, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the uninsured population variable in the model, it will cause the log of the percentage of black people coefficient to tend towards 0, or become less significant.

**Human Mobility Change in Transit Stations**

According to the results of the model a one unit increase in the mobility change in transit stations variable causes the rate of positive tests to increase by ~16 cases per 100,000 population. 

**Number of Days of Shelter in Place**

According to the results of the model a one unit increase in the number of shelter in place days reduces the case rate per 100,000 by ~0.8.

As previously discussed, States with densely-populated cities were the earliest to be impacted by the pandemic, and also subsequently the earliest to lock-down. Nevertheless, these States are also some of the wealthier ones due to being finance or business hubs, with higher levels of employment. We therefore predict a positive relationship between number of days of shelter in place and number of uninsured individuals in the population. 

Given that we predict a positive relationship between percentage of uninsured and the dependent variable (case rate per 100,000), and a positive relationship between percentage of uninsured percentage and the number of days of shelter in place, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the percentage of uninsured population percentage variable in the model, it will cause the shelter days coefficient to tend towards 0, or become less significant. As the coefficient is already negative, we would expect to see it becoming less negative.

**Number of Days of Non-Essential Businesses Closure**

According to the results of the model a one unit increase in the number of days that non-essential businesses remained closed causes the case rate to increase by ~6.9 cases per 100,000.

Unlike the shelter-in-place orders, the States that resisted closure of essential businesses and attempted to preserve their economies were predominantly those led by Republicans, which also happen to be States with higher levels of poverty (previously discussed and cited). As such we predict a negative relationship between number of days of enforced business closure and percentage population of uninsured individuals. 

Given that we predict a positive relationship between percentage of uninsured and the dependent variable (case rate per 100,000), and a negative relationship between percentage of uninsured percentage and the number of days of forced essential business closure, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be negative. That is, if we include the percentage of uninsured variable in the model, it will cause the business closure days coefficient to tend away from 0, or become more significant. 

## Population Density per Square Miles

This variable is included in the dataset.

Population density [varies by State](https://simple.wikipedia.org/wiki/List_of_U.S._states_by_population_density) across the United States. As expected, coastal States tend to have higher population densities, likely due to the presence of larger metropolitan areas within them. 

A larger population density may mean that more people are concentrated within a smaller area, which may then translate into increased opportunity for the spread of COVID-19. Moreover, inner-cities and low-income areas are likely to have higher population densities per unit of area. As they may be inhabited by socioeconomic groups and ethnicities that are at increased risk of being impacted by COVID-19, we predict that a larger population density is positively related to the test rate per 100,000 population. 

**Mandatory Mask Use**

According to this model, the enforcement of mask usage reduces the case rate by ~941 cases/100,000, holding the other variable constant.  

States hosting large cities are likely to be well aware of the risks of contracting COVID-19, especially if their cities have substantial areas of low-income housing or large concentrations of ethnic groups at increased risk of COVID-19. Subsequently they are more likely to impose mandatory mask-use policies. We predict that there is a positive relationship between mandatory mask use and population density.

Given that we predict a positive relationship between population density and the dependent variable (case rate per 100,000), and a positive relationship between population density and mask use enforcement, the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the population density variable in the model, it will cause the mask use coefficient to tend towards 0, or become less significant. As the coefficient is already negative, we would expect to see it becoming less negative.

**Test Rate per 100,000 habitants**

According to this model, increasing the test rate per 100,000 by 1 unit will increase the number of positive tests per 100,000 by 0.023.

It is likely that States with high population densities will be more proactive in early-diagnosis and control of COVID-19 spread. Moreover, if these States have higher rates of poverty or concentrations of racial minorities, we believe they will offer enhanced public health initiatives. As such, we predict a positive relationship between population density and test rate per 100,000 inhabitants. 

Given that we predict a positive relationship between population density and the dependent variable (case rate per 100,000) and a positive relationship between population density and test rate per 100,000 people, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the population density variable in the model, it will cause the test rate coefficient to tend towards 0, or become less significant.

**Percentage of Population Below 25 Years Old**

According to the results of the model, a one unit increase in the percentage of the population under the age of 25 increases the positive case rate by ~190 per 100,000 of the population, holding all other variables constant.

Black people comprise [31% and 24% of the US population between the ages 0-21 and 22-37](https://www.pewresearch.org/fact-tank/2019/07/30/most-common-age-among-us-racial-ethnic-groups/), respectively. We therefore predict a positive relationship between percentage of population under 25 and population density.

Given that we predict a positive relationship between the percentage of population density and the dependent variable (case rate per 100,000), and a positive relationship between percentage of population density and percentage of young people in the State, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the percentage population density variable in the model, it will cause the percentage of population coefficient to tend towards 0, or become less significant.

**Log of Percentage of Black Ethnicity in Total Population**

According to the results of the model, a one unit increase in the log of the percentage of black ethnicity in the population causes the rate of positive tests to increase by ~220 cases per 100,000 population. 

For reasons previously discussed, we predict that more densely populated areas are also likely to have higher rates of poverty or be located in inner-Cities. Subsequently, we predict a positive relationship between the total population density and percentage of Black people within the populations. 

Given that we predict a postive relationship between overall population density and the dependent variable (case rate per 100,000), and a positive relationship between the overall population density and the percentage of black people within a population, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the total population density variable in the model, it will cause the log of the percentage of black people coefficient to tend towards 0, or become less significant.

**Human Mobility Change in Transit Stations**

According to the results of the model a one unit increase in the mobility change in transit stations variable causes the rate of positive tests to increase by ~16 cases per 100,000 population. 

More highly populated areas are likely to be areas of business or commerce - i.e. more people commute to and from them daily. A large number of companies and businesses have implemented work-from-home initiatives, meaning fewer people need to commute to work daily. As a result we predict a negative relationship between population density and human mobility change in transit stations. 

Given that we predict a positive relationship between population density and the dependent variable (case rate per 100,000), and a negative relationship between population density and mobility change in transit stations, we predict that the overall predicted variable bias (disregarding all other independent variables) will be negative. That is, if we include the population density variable in the model, it will cause the human mobility change variable to tend away from zero, or become more significant.

**Number of Days of Shelter in Place**

According to the results of the model a one unit increase in the number of shelter in place days reduces the case rate per 100,000 by ~0.8.

As previously discussed, we predicted that local governments of areas with high population density were more likely to implement shelter-in-place orders earlier in the pandemic. We therefore predict that there is a positive relationship between population density and number of days of shelter in place. 

Given that we predict a positive relationship between the overall population density and the dependent variable (case rate per 100,000), and a positive relationship between overall population density and the number of days of shelter in place, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the percentage of uninsured population percentage variable in the model, it will cause the shelter days coefficient to tend towards 0, or become less significant. As the coefficient is already negative, we would expect to see it becoming less negative.

**Number of Days of Non-Essential Businesses Closure**

According to the results of the model a one unit increase in the number of days that non-essential businesses remained closed causes the case rate to increase by ~6.9 cases per 100,000.

Enforcement of shelter-in-place orders are likely to run concurrent with enforced non-essential business closures. We previously reasoned for the positive relationship between shelter in place orders and population density. Likewise, we believe there to be a positive relationship between number of days of non-essential business closure and State population density. 

Given that we predict a positive relationship between overall population density and the dependent variable (case rate per 100,000), and a positive relationship between overall population density and the number of days of forced essential business closure, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include the overall population density variable in the model, it will cause the business closure days coefficient to tend towards 0, or become less significant. 

## Statewide Education Levels

Under normal circumstances there is a positive relationship between [education level and the overall health of an individual](https://www.annualreviews.org/doi/abs/10.1146/annurev-publhealth-031816-044628). Though education is something of an multifactorial construct, ranging from basic literacy through to higher education, and therefore difficult to intrinsically measure, we predict that if it could be measured on a continuous scale (i.e. 0 being an adult who was completely illiterate, to 100 being an adult with an advanced post-graduate degree), there would be an negative relationship between educational attainment and positive test rates per 100,000 population. More educated populations are more likely to observe public health and good practices designed to reduce the spread of the disease. 

**Mandatory Mask Use**

According to this model, the enforcement of mask usage reduces the case rate by ~941 cases/100,000, holding the other variable constant.

States that have higher proportions of educated inhabitants are also likely to be ones that enforce mandatory public health measures like mask use. We therefore predict a positive relationship between education levels and mask use orders. 

Given that we predict a positive relationship between education levels and the dependent variable (case rate per 100,000), and a positive relationship between population education levels and mask use enforcement, the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include an education level variable in the model, it will cause the mask use coefficient to tend towards 0, or become less significant. As the coefficient is already negative, we would expect to see it becoming less negative.

**Test Rate per 100,000 habitants**

According to this model, increasing the test rate per 100,000 by 1 unit will increase the number of positive tests per 100,000 by 0.023.

As previously discussed, States with highly educated populations are also likely to have more robust healthcare infrastructures, including those offering access to COVID-19 testing. Consequently, we predict a positive relationship between test rate per 100,000 inhabitants and education levels within the population. 

Given that we predict a positive relationship between education levels and the dependent variable (case rate per 100,000) and a positive relationship between education levels and test rate per 100,000 people, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include an education level variable in the model, it will cause the test rate coefficient to tend towards 0, or become less significant.

**Percentage of Population Below 25 Years Old**

According to the results of the model, a one unit increase in the percentage of the population under the age of 25 increases the positive case rate by ~190 per 100,000 of the population, holding all other variables constant.

Relative to older populations, young people are more [likely to be educated](https://www.deseret.com/2015/8/10/20569870/today-s-young-adults-are-the-most-educated-group-in-history#pew-research-centers-fact-tank-recently-reported-that-todays-children-are-more-educated-than-their-grandparents-and-other-generations). As such, we predict a positive relationship between the percentage of the population under 25 years old and education levels. 

Given that we predict a positive relationship between education levels and the dependent variable (case rate per 100,000), and a positive relationship between percentage of education levels and percentage of young people in the State, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include an education level variable in the model, it will cause the percentage of population coefficient to tend towards 0, or become less significant.

**Log of Percentage of Black Ethnicity in Total Population**

According to the results of the model, a one unit increase in the log of the percentage of black ethnicity in the population causes the rate of positive tests to increase by ~220 cases per 100,000 population.

Due to longstanding socioeconomic reasons, there is an [observable gap](https://www.americanprogress.org/issues/education-postsecondary/reports/2018/05/23/451186/neglected-college-race-gap-racial-disparities-among-college-completers/) between educational attainment of minorities vs. non-minorities. As such, we predict a negative relationship between the percentage of black people within a population and educational levels in the population. 

Given that we predict a postive relationship between education levels and the dependent variable (case rate per 100,000), and a negative relationship between the education levels and the percentage of black people within a population, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be negative. That is, if we include the total population density variable in the model, it will cause the log of the percentage of black people coefficient to tend away from 0, or become more significant.

**Human Mobility Change in Transit Stations**

According to the results of the model a one unit increase in the mobility change in transit stations variable causes the rate of positive tests to increase by ~16 cases per 100,000 population. 

States with large metropolitan or commercial areas frequented by commuters were likely to have been affected substantially by work-from-home initiatives. As many commuters are likely to be more educated, we predict that states observing higher drops in mobility at transit stations are also those that have larger populations of highly educated people. There is therefore a negative relationship between these variables.  

Given that we predict a positive relationship between education levels and the dependent variable (case rate per 100,000), and a negative relationship between large populations of educated people and mobility change in transit stations, we predict that the overall predicted variable bias (disregarding all other independent variables) will be negative. That is, if we include the educational achievement variable in the model, it will cause the human mobility change variable to tend away from zero, or become more significant.

**Number of Days of Shelter in Place**

According to the results of the model a one unit increase in the number of shelter in place days reduces the case rate per 100,000 by ~0.8.

States that are proactive at implementing public health measures are more likely to have higher populations of highly educated people. We therefore predict a positive relationship between educational attainment and the number of shelter in place days. 

Given that we predict a positive relationship between the educational levels and the dependent variable (case rate per 100,000), and a positive relationship between educational levels of the population and the number of days of shelter in place, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include an educational level variable in the model, it will cause the shelter days coefficient to tend towards 0, or become less significant. As the coefficient is already negative, we would expect to see it becoming less negative.

**Number of Days of Non-Essential Businesses Closure**

According to the results of the model a one unit increase in the number of days that non-essential businesses remained closed causes the case rate to increase by ~6.9 cases per 100,000.

Similarly, shelter-in-place orders were likely concurrently issued together with enforced closure of non-essential businesses. For reasons discussed previously, we believe there is a positive relationship between shelter-in-place duration and educational achievement. Likewise, we also predict a positive relationship between number of days of non-essential business closure and the educational level of the population. 

Given that we predict a positive relationship between Statewide educational levels and the dependent variable (case rate per 100,000), and a positive relationship between Statewide educational levels and the number of days of forced essential business closure, we predict that the overall predicted omitted variable bias on the model (disregarding the other independent variables) will be positive. That is, if we include a Statewide educational level variable in the model, it will cause the business closure days coefficient to tend towards 0, or become less significant. 
